Modular Synthesis of Verified Verifiers of Computation with STV Algorithms

### 2019-01-26 paper/paper.tex
pushing the most up-to-date version

\documentclass[10pt,conference]{IEEEtran}
%\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{wrapfig}
\usepackage{amssymb,amsfonts}
%amsmath,
\usepackage{mathtools}

@@

\end{IEEEkeywords}
\section{Introduction}\label{sec:intro}

Single transferable vote (STV) \footnote{DP: I vaguelu reall that
there's a book that speaks about STV, at least in part? A reference
would be good here} is a preferential voting scheme used
in various jurisdictions around the world, including Ireland, Malta,
India, Pakistan, Australia, New Zealand and some municipialities of
the United States, typically for multi-seat constituencies. 
A ballot is a preference-ordered list of
candidates, typically obtained specifying a numerical preference
for each candidate appearing on the ballot paper as illustrated in Figure
\footnote{DP: remove figure and explain textually if low on space}
\ref{fig:ballot}.

\begin{figure}\label{fig:ballot}
\begin{center}
\includegraphics[width=0.18\textwidth]{bal-cropped.pdf}
\end{center}
\caption{A hypothetical STV ballot}
\end{figure}

Each specific STV scheme first determines a \emph{quota}, i.e. the
(minimal) number of votes that a candidate needs to attract to be
elected (often the Droop quota \footnote{DP: reference}). The quota
depends on the number of ballots cast, and the number of vacancies
to fill. Once the quota has been determined, tallying proceeds
as follows:
\begin{enumerate}
  \item count all first preference votes
  \item if a candidate's first preference votes are above the quota,
  this candidate is elected
  \item if an elected candidate has received votes in excess of
  the quota, these  excess votes are transferred to the next preference
  (possibly resulting in more candidates being elected)
  \item if there are still vacancies to fill and no candidate is
  being elected, the candidate with the least number of first
  preference votes is eliminated, and their votes are transferred to
  the next preference (possibly resulting in more candidates being
  elected)
  \item this process continues until either all vacancies are
  filled, or the number of elected candidates, together with the
  candidates not yet eliminated matches the number of vacancies.
\end{enumerate}

@@

We achieve this by first giving a formal definition STV (called
\emph{generic STV}) 
that 
encompasses the entire family of STV vote counting
schemes. This definition necessarily under-specifies any particular
STV vote counting scheme, but fixes both basic data structures and
mode of computation common to all instances of STV. Trustworthyness is achieved by
not verifying the \emph{software} that computes the result, but
instead by verifying a \emph{certificate} that substantiates the
correctness of the count. A certificate records the individual
steps taken in the computation of winners. (That is, our approach is
an instance of verifiable computation \footnote{DP: reference}). The
certificate directly corresponds to the individual steps of the
count (and could so also be verified, or spot-checked manually),
giving transparency. We minimize the computaional trust base by
verifying our executables down to the machine level using a
combination of HOL \footnote{DP: reference} and CakeML \footnote{DP:
reference}.

We observe that every instance of STV performs the same \emph{actions}
(e.g. count first preferences, transfer a surplus, eliminate a
candiate, etc.) where the precise interpretation of action differs.
We formalise these actions as rules that advance the state of the
count towards a final result. A certificate then becomes a
sequence of rule applications. While generic STV abstracts the
communalities, each \emph{instance} of STV formally specifies the
pre- and postconditions for each rule as well as a boolean function
that decides whether a rule has been applied correctly.

Executable certificate verifiers are the constructed almost
automatically: 
almost all of the ensuing proof obligations are discharged
generically, and we fully 
automate the subsequent translation into machine-level verified executable using CakeML.

@@

We evaluate our framework by analysing the degree of automation and
modularity achieved, and by giving realistic benchmarks results over
real-world eletion data.

In summary, we present a framework that allows us to formally
specify instances of STV and sythesise an executable
certificate verifier in a modular way that is itself verified down
to the machine level.
%Evaluation.
%\begin{itemize}
%  \item minimal trust base realised throug cakeml and hol
%  \item aplicability through benchmarks
%  \item adaptability: combinnation of rules are automatic
%  \item transparency through certificate format
%\end{itemize}

%\setlength{\abovedisplayskip}{0.5em}
%\setlength{\belowdisplayskip}{0.5em}
\section{Evidences}
\label{sec:DataEv}
%\section{The Generic STV Algorithm}
%\label{sec:GenSTV}
%%%%We shall initially clarify what the problem is that we are trying to address. 
%Then we proceed to elaborate on our approach and detail about the solution. 
%%%%%  To this end assume an arbitrary scheme $\mathcal{A}$ belongs to the STV family.The question is given any imp $\mathcal{P}$ is any implementation of $\mathcal{A}$ where the source code of $\mathcal{P}$ is  secret. 
%The question is how can one verify  $\mathcal{P}$ implements $\mathcal{A}$ correctly so that computation carried out by $\mathcal{P}$ is trustworthy?
%%%%The question is how can one verify that each execution of $\mathcal{P}$ correctly computes the end result in accordance with instructions in $\mathcal{A}$?

%Although full static verification is the most desirable answer to prove correctness, it is not practically possible here due to inaccessibility of source code. But even if the source were available it would not be feasible to carry such a task  because verifying every single implementation for any arbitrary STV algorithm is a Herculean significantly resource consuming agenda. Besides, any small change in implementation of an algorithm may render the whole  static verification of the implementation useless to the point that the process need to be redone from the scratch.   
We are motivated to 
%engineer a trustworthy practical  framework to answer the 
tackle the following problem. How can one verify that any execution of any implementation $\mathcal{P}$ whose source code is secret of an arbitrary  known STV algorithm $\mathcal{A}$ correctly computes the end result according to  $\mathcal{A}$?

To begin 
%with addressing the problem, 
assume $\mathcal{P}$ is an implementation of an algorithm $\mathcal{A}$. We reasonably demand each execution of $\mathcal{P}$ on a given input $x$ consisting of ballots recorded in the election to output the winners $y$ and evidence $\omega$ as claim  for correctness of the execution.
%of the election for an input $x$ consisting of election votes recorded. Given the situation, what is also sensible to ask and easy to be produce, is to make $\mathcal{P}$  also output an evidence $\omega$ of correctness of $y$. 
 Having evidence available for each execution of $\mathcal{P}$ facilitates checking correctness of the computation carried out \emph{independently} of the (source code of) $\mathcal{P}$. Therefore the original problem boils down to how one can verify the evidence of any \emph{instance of computation} with any algorithm $\mathcal{A}$.

%New questions emerge from the above problem reformulation.  
The evidence as such must be enough informative to provide transparency of tallying and also allow  voters, or at least a large pool of scrutineers, to verify for themselves that the tally is authentically processed as per instructions of the counting scheme  so that winners truly reflect will of the voters. 
%Note that evidences are produced by various instances of computation with many STV algorithms which differ from each other in their exact counting mechanism.  
Therefore the questions becomes what information is enough 
%for satisfying transparency and independent verifiability of tallying 
for establishing transparency and verifiability of tallying  and what data structure for recording the information in evidence we should choose.

%%As mentioned above, STV algorithms vary in their precise algorithmic content.  On the other hand, we aim at producing verified verifiers of evidences of these divergent algorithms through a modular process that allows us to synthesise evidence verifiers of different schemes  throughout a uniformly replicable procedure.  Hence another challenge is what generic mechanism can we craft to produce verified verifiers that validate evidences of an STV algorithm distinctively and yet be able to employ that same process for creating another verifier for a different scheme while minimising the effort in synthesising it.

\subsection{Characterisation of Evidence}
%%%Through thorough analysis of various STV schemes, we come to identify a common underlying data and algorithmic structure existing in the family. We abstract this structure as a finite state machine and call it the \emph{generic STV}\cite{}. Upon each execution on a given input $x$, the machine outputs a value $y$ and an \emph{evidence} $\omega$ which is the trace of all states that the machine visits from an initial machine state where $x$ is registered to the final state where $y$ is obtained.   
%%%%%Concrete STV algorithms are therefore realised as instantiations into the machine. Each of the instantiations  also produce an evidence for any instance of computation carried out. 
%%%%%%%Our objective here is to independently scrutinise process of counting under any STV schemes similar to physical scrutineers who observe tallying to attest its authenticity.  Having an evidence of computation at hand, we are then able to validate or reject the instance of computation carried out by checking if all of the transitions taken to terminate at the final state where the winners are declared are correct.  
A comprehensive analysis of different STV algorithms reveals existence of a common underlying abstract data type among STV algorithms as follows. 
%Each STV scheme operates on the following data.
\begin{itemize}
\item Some ballots to deal with (assuming) recorded as cast by voters. Each ballot is a pair consisting of a list expressing the preference of a voter and a fractional transfer value,
\item a quota which is the least amount of votes needed to attract in order to be elected, 
\item some candidates competing in the election, 
\item some number of vacancies for which candidates compete,
\item \emph{discrete states of computation} which encapsulate necessary  pieces of information needed to transparently know how the tallying has progressed from the beginning to the end and to be able to independently verify its correctness provided access to the information is granted. 
\end{itemize}
There are two kinds of discrete states of computation. Some of them are final states where tallying has reached a halting state by finding the winners of the election. The other consist of non-final ones where bits and pieces of data are manipulated according to counting mechanism of the STV scheme used, to eventually obtain an end result. Each non-final state comprises seven pieces of data: 
\begin{itemize}
 \item\textbf{uncounted ballots} which await being counted,
 \item\textbf{tally} for holding information about the amount of votes that each candidate has hitherto attracted %up to this state of computation,
 \item\textbf{pile} to know the ballots allocated to each candidate, 
 \item\textbf{backlog of elected} candidates who have exceeded the quota and their surplus votes awaits being transferred,
 \item\textbf{backlog of eliminated} ones whose votes awaits transfer,
 \item\textbf{elected candidates} who are the already elected ones whose surplus votes must be transferred,
 \item\textbf{continuing candidates} still continuing in the election. 
 \end{itemize}
%Data recorded in the final and non-final states as specified above   record all of the information needed and sufficient to know for independently validating a counting instance.

The non-final and final states specified above provide us with all information one needs. We therefore find a perfect characterisation for evidence. 
\begin{definition}\label{formalEv}
Suppose an STV algorithm $\mathcal{A}$ is given and $Q$, $s$ and $C$ are respectively the quota, number of vacancies and competing candidates in an electon whose counting algorithm is $\mathcal{A}$ . Also assume  $y$ is the output of an execution of $\mathcal{A}$ on an input $x$. By evidence $\hat{\omega}$ for the execution  we mean the quadruple $(Q,s,C,\Omega)$, where $\Omega$ is the chronological sequence of all states of this instance of computation visited from the initial state where $x$ is input to the final state where $y$ is output.   
\end{definition}
%\begin{definition}[Evidence]
%An evidence $\hat{\omega}$ for an instance of computation with an algorithm $\mathcal{A}$ is a trace of data whose abstract representation matches with some formal evidence. 
%\end{definition}
%Definition asserts the abstract data type whose values mathematically represent a state of computation. A sequence of these mathematical states formally represent an evidence. In other words, an evidence is a trace of data whose abstract characterisation is a member of the mathematical construct given in Definition. 
%\begin{definition}
%Assume $\mathit{Ballot}$, $\mathit{Tally}$, $\mathit{Pile}$ are respectively sets of ballots, tallies, and pile of ballots. Also assume $\mathcal{C}$ is the set of all of competing candidates in  an election. Then the set of final states equals to the set of all subsets of $C$, namely $2^{C}$, and the set of non-final states are 
%\begin{small}
%\begin{center}
%$\mathit{Ballot}\times\mathit{Tally}\times\mathit{Pile}\times (2^{C}\times 2^{C})\times 2^{C}\times 2^{C}$
%\end{center}
%\end{small}
%where the fourth component represents backlog of elected and removed and the last two ones are respectively for elected and continuing candidates. 
%\end{definition}
\section{The Generic STV  Machine}
\label{sec:Machine}
\setlength{\abovedisplayskip}{0.5em}
\setlength{\belowdisplayskip}{0.5em}
We have already discussed an abstract data type whose values  formally represent evidence. However we do not yet possess an operational semantics for manipulating data to validate evidence.   On one hand, the semantics must be flexible enough to accommodate variations existing in the counting mechanism of STV algorithms so that we can produce verifiers  specifically operating for validation of evidence output by their associated scheme. On the other hand it must facilitate designing and developing the framework modularly while offering automation of synthesising process of verifiers.
%in such a way that we achieve a desirable degree of modularity and automation of  the process of synthesising different verifiers.
%evidence verifiers for various STV algorithms.

%STV algorithms vary in their precise algorithmic content.  On the other hand, we aim at producing verified verifiers of evidences of these divergent algorithms through a modular process that allows us to synthesise evidence verifiers of different schemes  throughout a uniformly replicable procedure.  Hence another challenge is what generic mechanism can we craft to produce verified verifiers that validate evidences of an STV algorithm distinctively and yet be able to employ that same process for creating another verifier for a different scheme while minimising the effort in synthesising it.

%\subsection{STV Algorithms Analysis II}

@@

The underlying abstract data type and the universal algorithmic pattern found in STV characterise components of a finite state machine which we call \emph{generic STV}~(Definition\ref{STVMachine}). The set of machine states consist of the discrete states of computation $\mathcal{S}$ mentioned earlier. The seven generic actions form the transition labels $\mathcal{T}$ of the machine. Last, the everywhere present pre- and post-conditions which define (content of)  the actions, make a small-step operational semantics for the machine.
% \begin{remark} 
Note that the states of computation recorded in all possible  evidence now become identical with machine states.
%states of the generic STV. 
%\end{remark}

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ where  $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$. 
%where $S_t$ is the sanity check condition  for  transition $t \in \mathcal{T}$.  An \emph{instance} of $M$  is a tuple $I = \langle \mathcal{T}, (R_t)_{t \in \mathcal{T}} \rangle$, where for each $t \in \mathcal{T}$, $R_t \subseteq \mathcal{S} \times \mathcal{S}$ is a rule that satisfies the sanity check condition $S_t$. 
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
Differences exist between counting algorithms of STV schemes.  
 For example in the STV used for electing the upper house representatives of the Victoria state of Australia (Victoria STV), distributing votes of an eliminated candidate occurs step by step in order of the magnitude of the fractional values that those votes carry.  Consequently, a pre-condition for applying the action is to first assure that upon each instance of applying transfer-removed action (a) exactly those votes which have the same fractional value are transferred and (b) the transfer happens according to the magnitude order. In contrast, for example, the STV used in the lower house elections of Tasmania state of Australia (Tasmania STV) does not have such pre-condition simply because votes of an eliminated candidate are transferred in a single application of transfer-removed. There are also variations in post-conditions. For example transfer-elect of the ACT STV used in the lower house elections of the ACT state of Australia requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

@@

\begin{definition}[verifier]\label{verifier}
Assume $\hat{\mathcal{A}}= \langle \mathcal{S}, \mathcal{T}, (S'_t)_{t \in \mathcal{T}} \rangle$ is an instantiation of the STV algorithm $\mathcal{A}$. A verifier for instances of computation with $\mathcal{A}$ is a function $\hat{\mathcal{V}}$ mapping from $\mathbb{Q}\times\mathbb{N}\times 2^{C}\times$ \textsf{List}$(\mathcal{S})$ to $\{0,1\}$ such that for any evidence $\hat{\omega} = (Q,s,C,\Omega)$ where $\Omega=\langle\Omega_{1},\dots,\Omega_{n}\rangle$, $\hat{\mathcal{V}}(\hat{\omega}) = 1$ if and only if the following holds:\\
$\forall i\in\{1,\dots,n-1\}.$ 
%\hspace*{0.2cm}$\forall \Omega_{i-1}$ $\Omega_{i}\in\mathcal{S}.$
%$\exists \Omega_{1}$ $\Omega_{2}\in$ \textsf{List} $(\mathcal{S}).$ $\Omega = \Omega_{1} + [\Omega_{i-1};\Omega_{i}] + \Omega_{2}$  
$\exists t\in\mathcal{T}.$  $(\Omega_{i-1},\Omega_{i})\in t$ $\wedge~\vdash\mathcal{S}'_{t}[\Omega_{i-1},\Omega_{i}]$\\ 
where $\mathbb{Q}$ and $\mathbb{N}$ respectively represent the set of rational and natural numbers, $2^{C}$ is the set of all subsets of $C$, and \textsf{List}$(S)$ is the set of all possible lists of machine states. $\Omega_{i-1}$ is a pre-state and $\Omega_{i}$ is a post-state visited in the execution. The last line means that pre- and post-conditions  of applying $t$ to move from $\Omega_{i-1}$ to $\Omega_{i}$  are logically true. 
\end{definition}

We shall next provide an example to concretely illustrate a piece of evidence and exemplify how STV schemes work.

%In the rest of the paper we discuss technicalities of design  and development of the framework  for modular synthesis of formally verified verifiers. 
%However we first provide an example of an evidence to concretely illustrate them and exemplify how counting with an STV scheme works.

%Equipped with the suitable semantics, we can define what verifiers of evidences are and how they validate data provided in evidences. Assume a formal evidence $\omega$ $=$ $(Q,s,C,\Omega)$ for an instance of computation with an STV scheme $\mathcal{A}$ is given. A formal verifier $\mathcal{V}$ for $\mathcal{A}$ is a function that recursively checks if any two consecutive machine states $s_{1}$ and $s_{2}$ in $\Omega$ each two consecutive machine states in the sequence $\Omega$  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{A Concrete Evidence of a Real STV Algorithm}
% To provide an sensible example of what a real STV algorithm resembles and how a concrete instance of an evidence is, we first present the algorithm used in ACT state of Australia for electing the representatives of the lower house. Then we given an example of an execution of the algorithm on a fictional election. 
%\subsection{STV Algorithm of ACT State Australia}
%\label{subsec:ana}
%We continue the quest by first analysing the whole STV family. Through this examination we obtain insight into what  data evidences should encapsulate and what structure is suitable for organising the data.
%Moreover, we realise a generic method which lays the foundation for  modular design and implementation of the framework for synthesising verifiers.

%The protocol dictates how to perform computations in two sets of instructions. It first gives the general steps to be taken whenever appropriate and then details conditions required for those steps to occur.
%\begin{description}
%\item[Step~1.]

%%\noindent\textbf{Step~1.} Count the first preference votes for each candidate.
%\item[Step~2.]

%\noindent\textbf{Step~2.} calculate the quota according to the following formula;\\
%(total number of valid votes / (number of vacancies) + 1) + 1
%\item[Step~3.]

%%%\noindent\textbf{Step~3.} Any candidate with votes equal to or greater than the quota is declared elected. If all vacancies have been filled, then the election finishes. Otherwise, if any candidate have more votes than the quota then go to step~4, and if no candidate has exceeded the quota then go to step~5.
%\begin{itemize}
%\item[$\bullet$] if all vacancies have been filled, then the election finishes.
%\item[$\bullet$] if all vacancies have not been filled, does any candidate have more votes than the quota?
%\begin{itemize}
%\item if yes, then go to step~4
%\item if not then go to step~5
%\end{itemize}
%\end{itemize}
%\item[Step~4.]

%%%\noindent\textbf{Step~4.} Distribute the successful candidate's surplus votes to continuing candidates according to the further preferences shown on the ballot papers by those voters.

%\item[Step~5.]
%%%%\noindent\textbf{Step~5.} If there are more continuing candidates than there are vacancies remaining unfilled, exclude the candidate with the fewest votes and distribute this candidate's votes to continuing candidates according to the further preferences shown by those voters. Calculate each continuing candidate's new total votes then go back to step~3. If the number of continuing candidates is equal to the number of vacancies remaining, all of those candidates are declared elected and the election is completed.
%\end{description}

%The protocol further elaborates on the constraints for and effects of application of counting, electing, transferring, or eliminating candidates;
%\begin{description}
%\item[Counting the first preferences.]

%%%%\noindent\textbf{Counting first preferences.}   Count the number of first preference 
%(or number \lq\lq~1\rq\rq) votes for each candidate.  
%After all the formal first preference votes are counted, the quota can be calculated.  
%%%Any candidate who has votes equal to or greater than the quota is now elected.
%\item[Transferring surplus votes from elected candidates.]
 %If a candidate receives a total number of votes equal to or greater than the quota, the candidate is elected. If an elected candidate has received an exact quota of votes, all of those votes are set aside and not counted further.

%%%%\noindent\textbf{Transferring surplus of elected.}  The value of the surplus votes gained by an elected candidate is passed on to other candidates according to the preferences indicated on ballot papers by the voters. If a candidate has received more than a quota 
%of first preference votes, all the ballot papers received by the candidate are distributed at a reduced value called a fractional transfer value. If a candidate has received more votes than the quota following a transfer of votes from another elected candidate or from an excluded candidate, only that 
%%%%all the ballots that the candidate received at the last count of preferences which resulted in a surplus of votes,  called \emph{last parcel} of ballots, are distributed to continuing candidates at a fractional transfer value.
%After the surplus votes from an elected candidate have been distributed, the total number of votes which each candidate has received is recalculated. 
%%%%Any other candidates that have votes equal to or greater than the quota are elected. If vacancies remain, the surplus votes of any newly elected candidate are now  distributed one by one.
%\item[The fractional transfer value.] 
%\noindent\textbf{Fractional transfer value.}
%%%%The fractional transfer value is calculated using the following formula:\\
%%%(number of surplus votes) / (total number of ballot papers with further preferences shown)
%\item[Excluded candidates.]

%%%\noindent\textbf{Elimination.} If vacancies remain to be filled after all surplus votes from elected candidates have been distributed, the process of excluding the lowest-scoring candidate begins. 
%\begin{itemize} 
%%%% The candidate with the smallest number of votes is "excluded" and his or her ballot papers are distributed to continuing candidates according to the preferences shown by the voters.
%%%% Ballots from excluded candidates are distributed at the value at which they were received by the excluded candidate. %Ballot papers received by the candidate as first preference votes have a value of \lq\lq~1\rq\rq~, while ballot papers received following the distribution of a surplus will have a fractional transfer value. 
%This will vary depending on the group of surplus votes from which they were received. 
%%%At each stage after ballot papers have been distributed from an excluded candidate, the total votes received by each continuing candidate are recalculated.
% to determine whether any candidate has received votes equal to or greater than the quota. 
%\end{itemize}

>>>>

\documentclass[10pt,conference]{IEEEtran}
%\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amssymb,amsfonts}
%amsmath,
\usepackage{mathtools}

@@

\end{IEEEkeywords}
\section{Introduction}\label{sec:intro}
%\setlength{\abovedisplayskip}{0.5em}
%\setlength{\belowdisplayskip}{0.5em}
Single transferable vote (STV)\cite{DFar} is a preferential voting scheme used in various jurisdictions around the world, including Ireland, Malta, India, Pakistan, Australia, New Zealand and some municipalities of the United States, typically for multi-seat constituencies.  A ballot is a preference-ordered list of candidates, typically obtained specifying a numerical preference for each candidate appearing on the ballot paper.
% as illustrated in Figure
%\footnote{DP: remove figure and explain textually if low on space}
%\ref{fig:ballot}.

%\begin{figure}\label{fig:ballot}
%\begin{center}
%\includegraphics[width=0.18\textwidth]{bal-cropped.pdf}
%\end{center}
%\caption{A hypothetical STV ballot}
%\end{figure}

Each specific STV scheme first determines a \emph{quota}, i.e. the (minimal) number of votes that a candidate needs to attract to be elected (often the Droop quota \cite{Droop:1881:MER}). The quota, mostly but not always, depends on the number of ballots cast, and the number of vacancies to fill. Once the quota has been determined, tallying proceeds as follows:
\begin{enumerate}
  \item count all first preference votes
  \item if a candidate's first preference votes are above the quota, this candidate is elected
  \item if an elected candidate has received votes in excess of the quota, the surplus votes are transferred to the next preference 
  %(possibly resulting in more candidates being elected)
  \item if there are still vacancies to fill and no candidate is  being elected, the candidate with the least number of first preference votes is eliminated, and their votes are transferred to the next preference (possibly resulting in more candidates being elected)
  \item the process continues until either all vacancies are filled, or the number of elected and continuing (not yet eliminated) candidates matches the number of vacancies.
  %candidates together with the candidates not yet eliminated matches the number of vacancies.
\end{enumerate}

@@

We achieve this by first giving a formal definition STV (called \emph{generic STV}) 
that  encompasses the entire family of STV vote counting
schemes. This definition necessarily under-specifies any particular STV vote counting scheme, but fixes both basic data structures and mode of computation common to all instances of STV. Trustworthiness is achieved by
not verifying the \emph{software} that computes the result, but instead by verifying \emph{evidence} that substantiates the correctness of the count. Evidence records and corresponds to the individual steps taken in the computation of winners 
%(That is, our approach is an instance of verifiable computation \footnote{DP: reference}). 
%The evidence directly corresponds to the individual steps of the count
(and could so also be verified, or spot-checked manually), giving transparency. We minimize the computational trust base by verifying our executables down to the machine level using a combination of HOL \cite{hol} and CakeML \cite{cake}.

We observe that every instance of STV performs the same \emph{actions} (e.g. count first preferences, transfer a surplus, eliminate a candidate, etc.) where the precise interpretation of action differs. We formalise these actions as rules that advance the state of the count towards a final result. Evidence for an instance of computation then becomes a sequence of rule applications applied to perform this computation. While generic STV abstracts the commonalities, each \emph{instance} of STV formally specifies the
pre- and postconditions for each rule as well as a boolean function that decides whether a rule has been applied correctly.

Executable evidence verifiers are constructed almost
automatically:  almost all of the ensuing proof obligations are discharged generically, and we fully  automate the subsequent translation into machine-level verified executable using CakeML.

@@

We evaluate our framework by analysing the degree of automation and modularity achieved, and by giving realistic benchmarks results over real-world election data.

In summary, we present a framework that allows us to formally specify instances of STV and sythesise an executable certificate verifier in a modular way that is itself verified down to the machine level.
\section{Evidences}
\label{sec:DataEv}

We are motivated to 
%engineer a trustworthy practical  framework to answer the 
tackle the following problem. How can one verify that any execution of any implementation $\mathcal{P}$ whose source code is secret of an arbitrary  known STV algorithm $\mathcal{A}$ correctly computes the end result according to  $\mathcal{A}$?

To begin 
%with addressing the problem, 
assume $\mathcal{P}$ is an implementation of an algorithm $\mathcal{A}$. We reasonably demand each execution of $\mathcal{P}$ on a given input $x$ consisting of ballots recorded in the election to output the winners $y$ and evidence $\omega$ as claim  for correctness of the execution.  
 Having evidence available for each execution of $\mathcal{P}$ facilitates checking correctness of the computation carried out \emph{independently} of the (source code of) $\mathcal{P}$. Therefore the original problem boils down to how one can verify the evidence of any \emph{instance of computation} with any algorithm $\mathcal{A}$.

The evidence as such must be enough informative to provide transparency of tallying and also allow  voters, or at least a large pool of scrutineers, to verify for themselves that the tally is authentically processed as per instructions of the counting scheme  so that winners truly reflect will of the voters.    
Therefore the questions becomes what information is enough  
for establishing transparency and verifiability of tallying  and what data structure for recording the information in evidence we should choose.
\subsection{Characterisation of Evidence} 
A comprehensive analysis of different STV algorithms reveals existence of a common underlying abstract data type among STV algorithms as follows. 
%Each STV scheme operates on the following data.
\item Some ballots to deal with (assuming) recorded as cast by voters. Each ballot is a pair consisting of a list expressing the preference of a voter and a fractional transfer value,
\item a quota which is the least amount of votes needed to attract in order to be elected, 
\item some candidates competing in the election, 
\item some number of vacancies for which candidates compete,
\item \emph{discrete states of computation} which encapsulate necessary  pieces of information needed to transparently know how the tallying has progressed from the beginning to the end and to be able to independently verify its correctness provided access to the information is granted. 
There are two kinds of discrete states of computation. Some of them are final states where tallying has reached a halting state by finding the winners of the election. The other consist of non-final ones where bits and pieces of data are manipulated according to counting mechanism of the STV scheme used, to eventually obtain an end result. Each non-final state comprises seven pieces of data: 
 \item\textbf{uncounted ballots} which await being counted,
 \item\textbf{tally} for holding information about the amount of votes that each candidate has hitherto attracted %up to this state of computation,
 \item\textbf{pile} to know the ballots allocated to each candidate, 
 \item\textbf{backlog of elected} candidates who have exceeded the quota and their surplus votes awaits being transferred,
 \item\textbf{backlog of eliminated} ones whose votes awaits transfer,
 \item\textbf{elected candidates} who are the already elected ones whose surplus votes must be transferred,
 \item\textbf{continuing candidates} still continuing in the election. 
The non-final and final states specified above provide us with all information one needs. We therefore find a perfect characterisation for evidence. 
\begin{definition}\label{formalEv}
Suppose an STV algorithm $\mathcal{A}$ is given and $Q$, $s$ and $C$ are respectively the quota, number of vacancies and competing candidates in an electon whose counting algorithm is $\mathcal{A}$ . Also assume  $y$ is the output of an execution of $\mathcal{A}$ on an input $x$. By evidence $\hat{\omega}$ for the execution  we mean the quadruple $(Q,s,C,\Omega)$, where $\Omega$ is the chronological sequence of all states of this instance of computation visited from the initial state where $x$ is input to the final state where $y$ is output.   
\end{definition}
\section{The Generic STV  Machine}
\label{sec:Machine}
\setlength{\abovedisplayskip}{0.5em}
\setlength{\belowdisplayskip}{0.5em}
We have already discussed an abstract data type whose values  formally represent evidence. However we do not yet possess an operational semantics for manipulating data to validate evidence.   On one hand, the semantics must be flexible enough to accommodate variations existing in the counting mechanism of STV algorithms so that we can produce verifiers  specifically operating for validation of evidence output by their associated scheme. On the other hand it must facilitate designing and developing the framework modularly while offering automation of synthesising process of verifiers.

@@

The underlying abstract data type and the universal algorithmic pattern found in STV characterise components of a finite state machine which we call \emph{generic STV}~(Definition\ref{STVMachine}). The set of machine states consist of the discrete states of computation $\mathcal{S}$ mentioned earlier. The seven generic actions form the transition labels $\mathcal{T}$ of the machine. Last, the everywhere present pre- and post-conditions which define (content of)  the actions, make a small-step operational semantics for the machine.  
Note that the states of computation recorded in all possible  evidence now become identical with machine states.

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ where  $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$.  
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
Differences exist between counting algorithms of STV schemes.  
 For example in the STV used for electing the upper house representatives of the Victoria state of Australia (Victoria STV), distributing votes of an eliminated candidate occurs step by step in order of the magnitude of the fractional values that those votes carry.  Consequently, a pre-condition for applying the action is to first assure that upon each instance of applying transfer-removed action (a) exactly those votes which have the same fractional value are transferred and (b) the transfer happens according to the magnitude order. In contrast, for example, the STV used in the lower house elections of Tasmania state of Australia (Tasmania STV) does not have such pre-condition simply because votes of an eliminated candidate are transferred in a single application of transfer-removed. There are also variations in post-conditions. For example transfer-elect of the ACT STV used in the lower house elections of the ACT state of Australia requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

@@

\begin{definition}[verifier]\label{verifier}
Assume $\hat{\mathcal{A}}= \langle \mathcal{S}, \mathcal{T}, (S'_t)_{t \in \mathcal{T}} \rangle$ is an instantiation of the STV algorithm $\mathcal{A}$. A verifier for instances of computation with $\mathcal{A}$ is a function $\hat{\mathcal{V}}$ mapping from $\mathbb{Q}\times\mathbb{N}\times 2^{C}\times$ \textsf{List}$(\mathcal{S})$ to $\{0,1\}$ such that for any evidence $\hat{\omega} = (Q,s,C,\Omega)$ where $\Omega=\langle\Omega_{1},\dots,\Omega_{n}\rangle$, $\hat{\mathcal{V}}(\hat{\omega}) = 1$ if and only if the following holds:\\
$\forall i\in\{1,\dots,n-1\}.$ 
%\hspace*{0.2cm}$\forall \Omega_{i-1}$ $\Omega_{i}\in\mathcal{S}.$
%$\exists \Omega_{1}$ $\Omega_{2}\in$ \textsf{List} $(\mathcal{S}).$ $\Omega = \Omega_{1} + [\Omega_{i-1};\Omega_{i}] + \Omega_{2}$  
$\exists t\in\mathcal{T}.$  $(\Omega_{i-1},\Omega_{i})\in t$ $\wedge~\vdash\mathcal{S}'_{t}[\Omega_{i-1},\Omega_{i}]$\\ 
where $\mathbb{Q}$ and $\mathbb{N}$ respectively represent the set of rational and natural numbers, $2^{C}$ is the set of all subsets of $C$, and \textsf{List}$(S)$ is the set of all possible lists of machine states. $\Omega_{i-1}$ is a pre-state and $\Omega_{i}$ is a post-state visited in the execution. The last line means that pre- and post-conditions  of applying $t$ to move from $\Omega_{i-1}$ to $\Omega_{i}$  are logically true. 
\end{definition} 
We shall next provide an example to concretely illustrate a piece of evidence and exemplify how STV schemes work. 
 \subsection{An Example of a Concrete Piece of Evidence}
\label{subsec:InstEv}
Figure~\ref{EvInst} depicts evidence of a small election whose counting algorithm is the ACT STV 
with three candidates $a$, $b$ and $c$, and initially recorded ballots containing
$b_1=([a,c],1/1)$, $b_2=([a,b,c],1/1)$, $b_3=([a,c,b],1/1)$,
$b_4=([b,a],1/1)$, $b_5=([c,b,a],1/1)$.  The first component of each ballot is a preference-ordered list of candidates, and the second is the fractional transfer value (initially set to $1/1$). The evidence consists of a \emph{header} that specifies the quota (computed according to the ACT STV instructions), the number of vacancies, and the list of competing candidates. 
The fourth line is the election result, and the remainder of the evidence are non-final states visited to compute this
outcome. Each non-final state shows uncounted ballots, tallies of candidates, piles of votes attracted, backlog of the elected, backlog of the eliminated, list of elected candidates, and list of continuing candidates.
